Bayesian Decision Theory in Structural Geological Modeling - \\ How Reducing Uncertainties Affects Reservoir Value Estimations

### 2018-04-06 Risk analysis/Poster/Poster STC 2017/poster_landscape_sfont.tex
Changed figures from PNG to PDF
Small corrections poster

\begin{myblock}{2.1 - Methods}
\textbf{Construction of a 3D structural geological model:}
	\begin{itemize}
	%\item Model creation using GemPy (see \citet{gmd-2018-61})
	\item Synthetic model of a simple anticline-fault trap in a potential hydrocarbon system (see Figure \ref{fig:unc_lik})
	\item Inclusion of uncertainties by assigning probability distributions to the positions of layer interfaces in depth
	\end{itemize}							
								\vspace{0.5em}
								\begin{figure}
									\begin{minipage}{0.95\textwidth}
										\centering\includegraphics[width=1\textwidth]{figures/Uncertainties_Likelihoods}
										\caption{The structural geological model illustrated as a 2D cross section (A) and a 3D voxel representation (B). The inclusion of $z$-positional uncertainties affecting layer depths and fault offset are depicted in (C) and (D). Thicknesses of the three middle layers are defined by the distances of interface points (E).}
										\label{fig:unc_lik}
									\end{minipage}
								\end{figure}
\textbf{Integration in a probabilistic modeling framework for Bayesian analysis (see Figure \ref{fig:model_network}):}
	\begin{itemize}
	%\item Utilizing PyMC (see \citet{salvatier2016pymc3})
	\item Setting up a full probability model taking into account all parameter probability distributions
	\item Bayesian inference: Conditioning of parameters on additionally observed data via likelihood functions
	\item Approximation of posterior distributions through the use of MCMC sampling
	\end{itemize}
\textbf{Evaluation of modeling results:}
										\begin{itemize}
										\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties})
										\item Implementation of customized algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
										\item Decision making based on the optimization of a case-specific loss function that considers differently risk-affine actors
										\end{itemize}

@@

								\begin{minipage}[h]{0.598\textwidth} %0.656
								\textbf{Design of the custom loss function:}
								%We regard decision making as allocation of resources based on estimation. 
								We base value estimation on the 
								\begin{equation}\label{eq:LFR_final}
																L(\theta,\hat{\theta}) =
																\begin{cases}
																|\theta - \hat{\theta}|*r^{-0.5}, & \text{for } 0<\hat{\theta}<\theta  \\
																|\theta-\hat{\theta}|*a*r, & \text{for } 0<\theta<\hat{\theta} \\
																|\theta-\hat{\theta}|*b*r, & \text{for } \theta\leq0<\hat{\theta} \\
																|\theta-\hat{\theta}|*c*r^{-0.5}, & \text{for } \hat{\theta}\leq0<\theta 
																\end{cases},
																\text{ with } a,b,c,r \in \mathbb{Q}.
																\end{equation}
								\end{minipage}
								%\vspace{0.5em}
								\begin{minipage}{0.398\textwidth}
								\begin{figure}
									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H}
									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
									to be positioned in the	footwall (2). The maximum trap fill is defined by either the anticlinal spill point (3;a) or a point of leakage across the fault, depending on juxtapositions with layers
									underlying (3;b) or overlying the seal (3;c). The latter is only relevant if the
									critical Shale Smear Factor (SSF$_c$) is exceeded, as determined over D and T in (4).}
									\label{fig:trap_cond}
									%Trap closure is defined by the seal shape and the normal fault (3).
								\end{figure}
								%\begin{figure}
								%	\centering
								%	\includegraphics[width=1\textwidth]{figures/model_network}
								%	\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
								%	priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
								%	red. Arrows indicate direct connections from parent to child nodes.}
								%	\label{fig:model_network}
								%\end{figure}
								\end{minipage}
								%\begin{minipage}{0.327\textwidth}

@@

\begin{myblock}{3.1 - Results}				
	\vspace{0.2em}
	\begin{figure}
		\begin{minipage}{0.98\textwidth}
			%\hfill
			\begin{minipage}[t]{0.49\textwidth}
				\centering\includegraphics[width=0.99\textwidth]{figures/ML4}
				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV. Shannon entropy was significantly reduced in the posterior model (b). Cross-fault leakage to overlying layers was eliminated as a trap control mechanism (c) and coincides with the disappearance of the probability of failure in the posterior ROV distribution (d). The respective loss functions are plotted in (e), showing high decision convergence.}
				\label{fig:ML4}
			\end{minipage}
			%\hspace{length=0.1\textwidth}
			\begin{minipage}[t]{0.49\textwidth}
				\centering\includegraphics[width=0.99\textwidth]{figures/ML5}
				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
				\label{fig:ML5}
			\end{minipage}	
			%\hfill
		\end{minipage}
	\end{figure}

@@

\begin{myblock}{6 - Further Information}
This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017. \\
3D geological models were created using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework using PyMC \citep{salvatier2016pymc3}.

>>>>

\begin{myblock}{2.1 - Methods}
\textbf{Construction of a 3D structural geological model:}
	\begin{itemize}
	%\item Model creation using GemPy (see \citet{gmd-2018-61})
	\item Synthetic model of a simple anticline-fault trap in a potential hydrocarbon system (see Figure \ref{fig:unc_lik})
	\item Inclusion of uncertainties by assigning probability distributions to the positions of layer interfaces in depth
	\end{itemize}							
								\vspace{0.5em}
								\begin{figure}
									\begin{minipage}{0.95\textwidth}
										\centering\includegraphics[width=1\textwidth]{figures/Uncertainties_Likelihoods.pdf}
										\caption{The structural geological model illustrated as a 2D cross section (A) and a 3D voxel representation (B). The inclusion of $z$-positional uncertainties affecting layer depths and fault offset are depicted in (C) and (D). Thicknesses of the three middle layers are defined by the distances of interface points (E).}
										\label{fig:unc_lik}
									\end{minipage}
								\end{figure}
\textbf{Integration in a probabilistic modeling framework for Bayesian analysis:}
	\begin{itemize}
	%\item Utilizing PyMC (see \citet{salvatier2016pymc3})
	\item Setting up a full probability model taking into account all parameter probability distributions
	\item Bayesian inference: Conditioning of parameters on additionally observed data via likelihood functions
	\item Approximation of posterior distributions through the use of MCMC sampling
	\end{itemize}
\textbf{Evaluation of modeling results:}
										\begin{itemize}
										\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties})
										\item Implementation of customized algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
										\item Decision making based on the optimization of a case-specific loss function that considers differently risk-affine actors
										\end{itemize}

@@

								\begin{minipage}[h]{0.598\textwidth} %0.656
								\textbf{Design of the custom loss function:}
								%We regard decision making as allocation of resources based on estimation. 
								We base value estimation on the 
								\begin{equation}\label{eq:LFR_final}
																L(\theta,\hat{\theta}) =
																\begin{cases}
																|\theta - \hat{\theta}|*r^{-0.5}, & \text{for } 0<\hat{\theta}<\theta  \\
																|\theta-\hat{\theta}|*a*r, & \text{for } 0<\theta<\hat{\theta} \\
																|\theta-\hat{\theta}|*b*r, & \text{for } \theta\leq0<\hat{\theta} \\
																|\theta-\hat{\theta}|*c*r^{-0.5}, & \text{for } \hat{\theta}\leq0<\theta 
																\end{cases},
																\text{ with } a,b,c,r \in \mathbb{Q}.
																\end{equation}
								\end{minipage}
								%\vspace{0.5em}
								\begin{minipage}{0.398\textwidth}
								\begin{figure}
									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H.pdf}
									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
									to be positioned in the	footwall (2). The maximum trap fill is defined by either the anticlinal spill point (3;a) or a point of leakage across the fault, depending on juxtapositions with layers
									underlying (3;b) or overlying the seal (3;c). The latter is only relevant if the
									critical Shale Smear Factor (SSF$_c$) is exceeded, as determined over D and T in (4).}
									\label{fig:trap_cond}
									%Trap closure is defined by the seal shape and the normal fault (3).
								\end{figure}
								%\begin{figure}
								%	\centering
								%	\includegraphics[width=1\textwidth]{figures/model_network}
								%	\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
								%	priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
								%	red. Arrows indicate direct connections from parent to child nodes.}
								%	\label{fig:model_network}
								%\end{figure}
								\end{minipage}
								%\begin{minipage}{0.327\textwidth}

@@

\begin{myblock}{3.1 - Results}				
	\vspace{0.2em}
	\begin{figure}
		\begin{minipage}{0.98\textwidth}
			%\hfill
			\begin{minipage}[t]{0.49\textwidth}
				\centering\includegraphics[width=0.99\textwidth]{figures/ML4.pdf}
				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV. Shannon entropy was significantly reduced in the posterior model (b). Cross-fault leakage to overlying layers was eliminated as a trap control mechanism (c) and coincides with the disappearance of the probability of failure in the posterior ROV distribution (d). The respective loss functions are plotted in (e), showing high decision convergence.}
				\label{fig:ML4}
			\end{minipage}
			%\hspace{length=0.1\textwidth}
			\begin{minipage}[t]{0.49\textwidth}
				\centering\includegraphics[width=0.99\textwidth]{figures/ML5.pdf}
				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
				\label{fig:ML5}
			\end{minipage}	
			%\hfill
		\end{minipage}
	\end{figure}

@@

\begin{myblock}{6 - Further Information}
This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017 (see \citet{stamm2017}). \\
3D geological models were created using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework using PyMC \citep{salvatier2016pymc3}.
